How to use flowCore package

### 2006-12-11 inst/doc/HowTo-flowCore.Rnw
Work on vignette
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@21422 bc3139a8-67e5-0310-9ffc-ced21a209358

\section{Methodology}
\subsection{Reading data}
We first load the libraries and read the data into R as FCS object 
(rflowcyt data format). Note than this small example works for cytoFrame object. 
<<load library>>=
library("flowCore")
@ 
<<load data,echo=FALSE,results=hide>>=
fcs.loc <- system.file("extdata", package="flowCore")
file.location <- paste(fcs.loc, dir(fcs.loc), sep="/")
b08 <- read.FCS(file.location[1])
e07 <- read.FCS(file.location[2])
f06 <- read.FCS(file.location[3])
@
\begin{verbatim}
> fcs.loc <- system.file("extdata", package="flowCore")
> file.location <- paste(fcs.loc, dir(fcs.loc), sep="/")
> b08 = read.FCS(file.location[1])
> e07 = read.FCS(file.location[2])
> f06 = read.FCS(file.location[3])
\end{verbatim}

@@

<<plotTRansfo,eval=FALSE, echo=FALSE,fig=TRUE,prefix=FALSE,include=FALSE>>=
##we should find a way to specified if an parameters has been transformed 
## either in the title or via the label of the axis 
plot(data08LogTransform,plotParameters=c("FSC-H","SSC-H"))
@

\subsection{Filtering data (gating)}
\subsubsection{Rectangular gate}
We will define rectangular gates and apply them to filter the bO8 data.
Figure \ref{fig:Fig02} presents the results of such filtering. 
The plot in Figure \ref{fig:Fig02}A shows the result of 
filtering out the data points that are below 200 in the FSC-H channel.
Figure \ref{fig:Fig02}B shows the result of 
filtering out the data points are between 200 
and 500  in the FSC-H channel.

<<define and apply rectangleGate1,eval=TRUE, echo=FALSE,results=hide>>=
min1 <- 200
max1 <- Inf
fil1 <- new("rectangleGate",filterId="nonDebris",
            parameters="FSC-H",min=min1,max=max1)
b08fil1 <- applyFilter(fil1,b08)
@ 
<<plotfiltered, eval=TRUE, echo=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(b08,y=b08fil1,pch="*")
@ 
<<define and apply polygonGate2,eval=TRUE, echo=FALSE,results=hide>>=
min2 <- 200
max2 <- 500
centerFilter <- new("rectangleGate",filterId="center",
            parameters=c("FSC-H"), min=min2, max=max2)
b08centerFilter <- applyFilter(centerFilter,b08)
@ 
<<plotCenterfiltered, eval=TRUE, echo=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(b08,y=b08centerFilter,pch="*")
@

@@

\subsubsection{polygon gate}
%%We will define a rectangular gate and apply the filter it to the bO8 data.
<<define and apply rectangleGate,eval=TRUE,echo=FALSE>>=
sqrcut <- matrix(c(300,300,600,600,50,300,300,50),ncol=2,nrow=4)
sqrfilter <- new("polygonGate",filterId="nonDebris",
            parameters=c("FSC-H","SSC-H"),boundaries= sqrcut)
b08sqrfilter <- applyFilter(sqrfilter,b08)
@

@@

<<define and apply polygonGate,eval=TRUE,echo=FALSE>>=
triangleCut <- matrix(c(300,400,600,300,400,300),ncol=2,nrow=3)
trianglefilter <- new("polygonGate",filterId="nonDebris",
            parameters=c("FSC-H","SSC-H"),boundaries= triangleCut)
b08Trgfilter <- applyFilter(trianglefilter,b08)
@

@@

The first gate is a range gate to filter out debris

<<SelectNonDebris,eval=TRUE,echo=FALSE,results=hide>>=
min1 <- c(200,0)
max1 <- c(800,800)
filter1 <- new("rectangleGate",filterId="Nondebris",parameters=c("FSC-H","SSC-H"),
  min=min1,max=max1)

b08.result1 <-  applyFilter(filter1,b08)
e07.result1 <- applyFilter(filter1,e07)
f06.result1 <- applyFilter(filter1,f06)

tab <- c("B08"=sum(b08.result1@subSet),"E07"=sum(e07.result1@subSet),"F06"=sum(f06.result1@subSet))
@ 
<<plotResFilter1, echo=FALSE, eval=FALSE,  fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(b08,y=b08.result1,main="B08",pch="*")
@ 
<<plotResFilter2, echo=FALSE, eval=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(e07,y=e07.result1,main="E07",pch="*")
@ 
<<plotResFilter3, echo=FALSE, eval=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(f06,y=f06.result1,main="F06",pch="*")
@ 
%%\begin{figure}
%%\centering
%%\subfigure{\includegraphics[width=40mm]{plotResFilter1}}
%%\subfigure{\includegraphics[width=40mm]{plotResFilter2}} 
%%\subfigure{\includegraphics[width=40mm]{plotResFilter3}} 
%%\caption{
%%  \label{fig:Fig05}{Example of the effect of filtering with a
%%    polygon gate.}} 
%%\end{figure}

@@

<<FilterLiveCell,eval=TRUE,echo=FALSE,results=hide>>=
filter2 = new("norm2Filter",filterId="Live Cells",scale.factor=2,method="covMcd",parameters=c("FSC-H","SSC-H"))

b08.result2 = applyFilter(filter2,b08,parent=b08.result1)
e07.result2 = applyFilter(filter2,e07,parent=e07.result1)
f06.result2 = applyFilter(filter2,f06,parent=f06.result1)

tab <- rbind(tab,c(sum(b08.result2@subSet),sum(e07.result2@subSet),sum(f06.result2@subSet)))

@@

<<FL1Filter,eval=TRUE,echo=FALSE,results=hide>>=
filter3 <- new("rectangleGate",filterId="FL1-H+",parameters="FL1-H",min=400,max=Inf)
b08.result3 <- applyFilter(filter3,b08,b08.result2)

filter4 <-new("norm2Filter",filterId="FL1-H+",scale.factor=2,method="covMcd",parameters=c("FSC-H","FL1-H"))
b08.result4 <- applyFilter(filter4,b08,b08.result2)

>>>>

\section{Methodology}
\subsection{Reading data}
We first load the libraries and read the data into R as FCS object 
(rflowcyt data format). Note than this small example works for cytoFrame object. 
<<load library>>=
library("flowCore")
@ 
<<load data,echo=FALSE,results=hide>>=
fcs.loc <- system.file("extdata", package="flowCore")
file.location <- paste(fcs.loc, dir(fcs.loc), sep="/")
b08 <- read.FCS(file.location[1],transformation="scale")
e07 <- read.FCS(file.location[2],transformation="scale")
f06 <- read.FCS(file.location[3],transformation="scale")
@
\begin{verbatim}
> fcs.loc <- system.file("extdata", package="flowCore")
> file.location <- paste(fcs.loc, dir(fcs.loc), sep="/")
> b08 = read.FCS(file.location[1])
> e07 = read.FCS(file.location[2])
> f06 = read.FCS(file.location[3])
\end{verbatim}

@@

<<plotTRansfo,eval=FALSE, echo=FALSE,fig=TRUE,prefix=FALSE,include=FALSE>>=
##we should find a way to specified if an parameters has been transformed 
## either in the title or via the label of the axis 
plot(data08LogTransform,plotParameters=c("FSC-H","SSC-H"),main="Log2 transformed data")
@

\subsection{Filtering data (gating)}
\subsubsection{Rectangular gate}
We will define rectangular gates and apply them to filter the bO8 data.
Figure \ref{fig:Fig02} presents the results of such filtering. 
The plot in Figure \ref{fig:Fig02}A shows the result of 
filtering out the data points that are below 0.2 in the FSC-H channel.
Figure \ref{fig:Fig02}B shows the result of 
filtering out the debris that are between 0.2 and 0.8 in the FSC-H channel and 
and 0 and 0.8 in the SSC-H channel.

<<define and apply rangeGate1,eval=TRUE, echo=FALSE,results=hide>>=
rangeGate <- rectangleGate(filterId="rangeGate","FSC-H"=c(.2,Inf))
b08rangeGate <- filter(b08,rangeGate)
@ 
<<plotfiltered, eval=TRUE, echo=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(b08,y=b08rangeGate, pch="*",main="B08, lower debris")
@ 
<<define and apply ploygoneGate2,eval=TRUE, echo=FALSE,results=hide>>=
rectGate <- rectangleGate(filterId="nonDebris","FSC-H"=c(0.2,0.8),"SSC-H"=c(0,0.8))
b08rectGate <- filter(b08,rectGate)
b08Filtered <- b08[b08rectGate,]
@ 
<<plotCenterfiltered, eval=TRUE, echo=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(b08,y=b08rectGate,pch="*",main="B08 - Nondebris")
@

@@

\subsubsection{polygon gate}
%%We will define a rectangular gate and apply the filter it to the bO8 data.
<<define and apply rectangleGate,eval=TRUE,echo=FALSE>>=
##sqrcut <- matrix(c(300,300,600,600,50,300,300,50),ncol=2,nrow=4)
##sqrfilter <- new("polygonGate",filterId="nonDebris",
##            parameters=c("FSC-H","SSC-H"),boundaries= sqrcut)
sqrfilter <- polygonGate(filterId="nonDebris","FSC-H"=c(0.3,0.3,0.6,0.6),"SSC-H"=c(0.05,0.3,0.3,.05))
b08sqrfilter <- filter(b08,sqrfilter)
@

@@

<<define and apply polygonGate,eval=TRUE,echo=FALSE>>=
triangleCut <- matrix(c(0.3,0.4,0.6,0.3,0.4,0.3),ncol=2,nrow=3)
colnames(triangleCut) <- c("FSC-H","SSC-H")
trianglefilter <- polygonGate(filterId="nonDebris",boundaries= triangleCut)
b08Trgfilter <- filter(b08,trianglefilter)
@

@@

The first filter is a serie of rectangle gates to filter out debris

<<SelectNonDebris,eval=TRUE,echo=FALSE,results=hide>>=
filter1 <- rectangleGate(filterId="nonDebris","FSC-H"=c(0.2,0.8),"SSC-H"=c(0,0.8))
b08.result1 <-  filter(b08,filter1)
##e07.result1 <- filter(e07,filter1)
##f06.result1 <- filter(f06,filter1)
summary(b08.result1)
##tab <- c("B08"=sum(b08.result1@subSet),"E07"=sum(e07.result1@subSet),"F06"=sum(f06.result1@subSet))
@ 
<<plotResFilter1, echo=FALSE, eval=FALSE,  fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(b08,y=b08.result1,main="B08 nonDebris",pch="*")
@ 
<<plotResFilter2, echo=FALSE, eval=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(e07,y=e07.result1,main="E07 nonDebris",pch="*")
@ 
<<plotResFilter3, echo=FALSE, eval=FALSE, fig=TRUE, prefix=FALSE, include=FALSE>>=
plot(f06,y=f06.result1,main="F06 nonDebris",pch="*")
@ 
%%\begin{figure}
%%\centering
%%\subfigure{\includegraphics[width=40mm]{plotResFilter1}}
%%\subfigure{\includegraphics[width=40mm]{plotResFilter2}} 
%%\subfigure{\includegraphics[width=40mm]{plotResFilter3}} 
%%\caption{
%%  \label{fig:Fig05}{Example of the effect of filtering with a
%%    polygon gate.}} 
%%\end{figure}

@@

<<FilterLiveCell,eval=TRUE,echo=FALSE,results=hide>>=
##filter2 <- norm2Filter(filterId="Live Cells",scale.factor=2,method="covMcd",parameters=c("FSC-H","SSC-H"))
filter2 <- norm2Filter("FSC-H","SSC-H",filterId="Live Cells",scale.factor=2,method="covMcd")
b08.result2 <-  filter(b08, filter2 %subset% b08.result1)

##e07.result2 <- filter(e07,filter2,parent=e07.result1)
##f06.result2 <- filter(f06,filter2,parent=f06.result1)

summary(b08.result2)
##tab <- rbind(tab,c(sum(b08.result2@subSet),sum(e07.result2@subSet),sum(f06.result2@subSet)))

@@

<<FL1Filter,eval=TRUE,echo=FALSE,results=hide>>=
filter3 <- rectangleGate(filterId="FL1-H+","FL1-H"=c(400,Inf))
b08.result3 <- filter(b08,filter3 %subset%b08.result2)

filter4 <- norm2Filter("FSC-H","FL1-H",filterId="FL1-H+",scale.factor=2,method="covMcd")
b08.result4 <- filter(b08,filter4 %subset% b08.result2)
